---
title: "Exploring Abalone Dataset"
author: "Javier Moreno"
date: "18 de febrero de 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Source: https://code.datasciencedojo.com/datasciencedojo/datasets/tree/master/Abalone

## Exploring Abalone Dataset

### Data Dictorionary
Get more information on the source.
Basically, rings +1.5 give the age in years

##Importing the dataset from the .csv
First of all, let's check the working directory and go to Documents:
```{r}
getwd()
```

Let's move to "Documents" folder:
```{r}
setwd("C:/Users/jmoreno7/Documents")
```

If we can't remember the name of the .csv, we can use RegEx:
```{r}
list.files(".", pattern="*.csv")
```

We will use 'tidyverse' for exploring data:
```{r}
library(tidyverse)
```

For importing data, we will use the `read_csv` function from tidyverse, althoguh `read.delim` or `read.csv` from base are also useful.

```{r}
dataset <- read_csv("Abalone.csv")
```

Let's do a first approach to the dataset

```{r}
#Let's cheack the heads
head(dataset)
#Let's check the data type
str(datase)
```

The dataset doesn't contain header, so let's add it manually:
```{r}
dataset <- read_csv("Abalone.csv",col_names = c("sex","length","diameter", "height","whole_weight","shucked_Weight","viscera","shell","rings"))
```

Taking  alook on the dataset, variable 'sex' should be a factor. Should be changed:
```{r}
dataset$sex <- as_factor(dataset$sex)
```

Note that "I" means Infant

We know that the age is the number of rings +1.5:

```{r}
df = dataset %>% mutate(age = rings + 1.5)
```

Our dataset has this dimensions:

```{r}
dim(df)
```

¿Do it have NaN values?

```{r}
summary(is.na.data.frame(df))
```

There are no NaN

Let's get a first approach to the dataset:
```{r}
summary(df)
```

We are going to do a correlation matrix:

```{r}
dfcor <- cor(df, method="pearson")
```

We get an error because there are some values which are not numerical: sex, which is a factor. Let's remove sex.

```{r}
df_cor <- df %>% select(-sex)
dfcor <- round(cor(df_cor, method = "pearson"), digits=2)
```

Let's take a look on the correlation numbers:

```{r}
dfcor
```

This is difficult to see. We will use the library `corrplot`
```{r}
library(corrplot)
corrplot(dfcor)
```

Or in order to see it clearer, we can use this configuration:
```{r}
library(corrplot)
corrplot(dfcor, method="number", type="upper")
```

```{r}
library(PerformanceAnalytics)
chart.Correlation(dfcor, histogram=TRUE)
```

We can take a lok about how sex affects on rings
```{r}
library(ggplot2)
df %>% ggplot(aes(rings, color = sex)) +
  geom_density()
 
```

```{r}
df %>% ggplot(aes(x=sex, y=rings, color=sex)) +
  geom_boxplot()
 
```

We can see there's no difference between number of rings and sex, despite the "infants"

We can get information about the different groups using `purrr`, `split` by sex and `summary`.
```{r}
library(purrr)
df %>% split(.$sex) %>% map(summary)
```

It shows there are odd data on Infants, showing that there are "Infants" woth an age over 22...

```{r}
df %>% ggplot(aes(x=age, y=length, col=sex)) +
  geom_point()+
  facet_grid(rows=vars(sex))
```







