---
title: "Slot Machine Simulation"
author: "Javier Moreno (Based on KSchool syllabus Example)"
date: "12 de febrero de 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Slot Machine Simulation

This example includes a simultaion of a slot machine and how much guves on prizes for each bet.

Let'start building the slot machine

```{r}
combinations <- c("DD","7","BBB","BB","B","C","*")
names(combinations) <- c("Diamonds","Seven","Triple Bars","Double Bars","Bars","Cherry","Others")

probabilities <- c(0.03,0.03,0.06,0.10,0.25,0.01,0.52)
sum(probabilities) == 1

df <- data.frame(figures=combinations,probs=probabilities,
                 stringsAsFactors = FALSE)

View(df)
```

### Strategy

Now, we will reduce the problem in different parts:

- Create a function called 'get_figures' which generates three figures based on given probabilities

- Get the prize of the combinations

### 'get_figures'

In order to get the figures, we create the function

```{r}
get_figures <- function(){
  result <- sample(df$figures, size=3, replace=TRUE, prob=df$probs)
  return(result)
}

```

Let's create a dummy version to evaluate the prizes

```{r}
my_figures <- get_figures()
```

```{r}
my_figures
```

### 'prizes'
We have to create a function which matches the result with the prizes. There are some questions:

- Are all of the results equal?

    - Different ways to answer this:
        - Comparing one by one
```{r}
set <- c("DD","DD","DD")
(set[1] == set[2]) & (set[1] == set[3])
```
        
        - Using `all`
```{r}
set <- c("DD","DD","DD")
all(set == set[1])
```
    
        - Using `length(unique)==1` 
```{r}
set <- c("DD","DD","DD")
unique(set)
length(unique(set))==1
```

- Are all the results part of 'B' family?
    -  We can check it
```{r}
set <- c("B","B","BBB")
all(set %in% c("B","BB","BBB"))

set2 <- c("B","*","BBB")
all(set2 %in% c("B","BB","BBB"))
```

- But careful: there's overlapping between this method and "B","B","B" and prices are different.

- How many cherries do we have?
```{r}
set <- c("B","C","BBB")
sum(set == "C")

set2 <- c("C","*","C")
sum(set2 == "C")
```

Now we can buil the 'prize' function:

```{r}
prize <- function(x){
  #Identify the case:
  equal <- length(unique(x))==1
  bars <- all(x %in% c("B","BB","BBB"))
  cherries <- sum(x=="C")
  
  #Prize calculation
  if (equal & x[1]!="*"){
    payout <- c("DD"=100,"7"=80,"BBB"=40,"BB"=25,"B"=10,"C"=10)
    prize <- unname(payout[x[1]])
  }else if(bars & !equal){
    prize <- 5
  }else if (cherries > 0){
    if (cherries == 2){
      prize <- 5
    } else {
      prize <- 2
    }
  }else{
    prize <- 0
  }
  return(prize)
}
```

And join both on 'play' function

```{r}
play <- function(){
  my_comb <- get_figures()
  my_prize <- prize(my_comb)
  print(my_comb)
  return(my_prize)
}

play()
```

Let's play:

```{r}
play()
```

## Loop to test the pay-back
Now we are going to create a loop to test how much money the slot machine gives back for each bet.

First of all, let's ging to generate a data.frame with all different combinations using `expand.grid`

```{r}
wheel <- c("DD","7","BBB","BB","B","C","*")
combos <- expand.grid(wheel_1=wheel, wheel_2=wheel, wheel_3=wheel, stringsAsFactors = FALSE)
head(combos)
```

Now we have set a data frame with 3 columns (wheel_1, whee_2, wheel_3).
We can assign the probabilities for each figure in each wheel.

```{r}
probabilities <- c("DD"=0.03,"7"=0.03,"BBB"=0.06,"BB"=0.10,"B"=0.25,"C"=0.01,"*"=0.52)
combos$prob_1 <- probabilities[combos$wheel_1]
combos$prob_2 <- probabilities[combos$wheel_2]
combos$prob_3 <- probabilities[combos$wheel_3]

head(combos)
```

Therefore, the total probability for each combination is:

```{r}
combos$totalprob <- combos$prob_1 * combos$prob_2 * combos$prob_3
head(combos)
```

If this is correct, the total sum of probabilities would be 1, which is the case.

```{r}
sum(combos$totalprob)
```

And the most probable combination is * * * with a probability of the 14,06%
```{r}
filter = max(combos$totalprob)
combos[combos$totalprob == filter,]
```

Now we have this, we are going to use a for...loop... to add the prize for each combination:

```{r}
for (k in 1:nrow(combos)){
  figures <- c(combos[k,1], combos[k,2], combos[k,3])
  combos$prize[k] <- prize(figures)
}
head(combos)
```

Instead of using a loop, it's also possible to make the calculations with `apply` function:

```{r}
score.combo <- function(x){prize(x[1:3])}
combos$prize <- apply(combos,1,score.combo)
```

Now we can calculate the payback:

```{r}
sum(combos$prize * combos$totalprob)
```

So the slot machine gives back approximately the 51.54% of the total bet.

